Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-17 journal paper/content/sec-intro.tex
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

While we build on a vast and active research area, we present sufficient background in Appendix~\ref{sec:related-classes} so as to keep this paper accessible to non-specialists.

>>>>

While we build on a vast and active research area, we present sufficient background in Appendix~\ref{app:bg} so as to keep this paper accessible to non-specialists.

### 2020-07-17 journal paper/content/sec-theory.tex
fix typos
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}\label{eq:theory-BIR-defn}
\BIR_i(n) := \E\sbr{ \max_{a\in A} \mu_a - \rew_i(n)}.
\end{align}
We are primarily interested in how fast \BIR decreases with $n$. (We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:examples} for background.

An algorithm is called \emph{monotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
Appendix~\ref{app:MAB-monotone}.

@@

\begin{corollary}\label{cor:DG-dominance}
The competition game has a unique Nash equilibirium: both principals choose \DynGreedy. \asedit{This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.}
\end{corollary}

>>>>

We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}\label{eq:theory-BIR-defn}
\BIR_i(n) := \E\sbr{ \max_{a\in A} \mu_a - \rew_i(n)}.
\end{align}
We are primarily interested in how fast \BIR decreases with $n$. (We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:bg} for background.

An algorithm is called \emph{monotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
Appendix~\ref{app:MAB-mono}.

@@

\begin{corollary}\label{cor:DG-dominance}
The competition game has a unique Nash equilibrium: both principals choose \DynGreedy. \asedit{This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.}
\end{corollary}

### 2020-07-17 journal paper/content/sim_details.tex
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model}, and look for qualitative differences between them. We take a representative algorithm from each class. Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
We use \Thompson (\TS) from the ``adaptive-exploration" algorithms, \DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,%
\footnote{In each round, \DynamicEpsGreedy algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data. We use $\eps=5\%$ throughout. Our pilot experiments show that choosing a different $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}. For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution. Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
Self-contained background can be found in Appendix~\ref{sec:related-classes}.

>>>>

\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model} (and in more detail in Appendix~\ref{app:bg}), and look for qualitative differences between them. We take a representative algorithm from each class. Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
We use \Thompson (\TS) from the ``adaptive-exploration" algorithms, \DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,%
\footnote{In each round, \DynamicEpsGreedy algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data. We use $\eps=5\%$ throughout. Our pilot experiments show that choosing a different $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}. For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution. Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.

%Self-contained background can be found in Appendix~\ref{sec:related-classes}.

### 2020-07-17 journal paper/itcs18paper/app-examples.tex
Minor edits.
Added comments.
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

\subsection{Monotonicity-in-information for the greedy step}
\label{sec:MAB-greedy}

\ascomment{TODO: insert an intro sentence.}

@@

\ascomment{TODO: find a citation for Lemma~\ref{lm:MII}.}

@@

\subsection{Monotone MAB algorithms}
\label{app:MAB-monotone}

\ascomment{Steven: NB: this is the point of this subsection...}
This subsection argues that ``monotonicity" is a mild assumption in our context. To this end, we present versions of some standard MAB algorithms that are monotone and satisfy ``standard" regret bounds.

@@

\begin{lemma}
Consider an algorithm which initially explores each action $m$ times (where $m$ is a parameter), chooses an action  with the highest observed average reward, and uses it for the remaining steps. In the initial exploration phase, actions are randomly permuted. 
\begin{itemize}
\item[(a)] This algorithm is monotone, for every \MRV.
\ascomment{Steven: right?}

\item[(b)] Fix $\delta>0$ and 
    $m = T^{2/3}\log(|A|/\delta)$.
Then, with probability $1-\delta$, it holds that  
    $\IReg(t)=\tildeO\rbr{T^{-1/3}}$
for each step $n> m\,|A|$.
\ascomment{Steven: can you derive a statement on regret, too?}
\end{itemize}
\end{lemma}

@@

\begin{lemma}
Posit $K=2$ arms and Bernoulli rewards. 
\begin{itemize}
\item[(a)] Algorithm \PhasedExplorExploit is monotone, for every \MRV.
    \ascomment{Steven: right?}

>>>>

\subsection{Monotone algorithms}
\label{app:MAB-mono}

%This subsection argues that ``monotonicity" is a mild assumption in our context. To this end, we present versions of some standard MAB algorithms that are monotone and satisfy ``standard" regret bounds.

\ascomment{TODO: insert an intro sentence, cite the ec15 paper.}

@@

@@

%\subsection{Monotone MAB algorithms}
%\label{app:MAB-monotone}

@@

\begin{lemma}
Consider an algorithm which initially explores each action $m$ times (where $m$ is a parameter), chooses an action  with the highest observed average reward, and uses it for the remaining steps. In the initial exploration phase, actions are randomly permuted.
\begin{itemize}
\item[(a)] This algorithm is monotone.
\ascomment{Steven: for every \MRV or only in expectation?}

\item[(b)] Fix $\delta>0$ and
    $m = T^{2/3}\log(|A|/\delta)$.
Then, with probability $1-\delta$, it holds that
    $\IReg(t)=\tildeO\rbr{T^{-1/3}}$
for each step $n> m\,|A|$.
\ascomment{Steven: can you derive $O(T^{2/3})$ regret, too?}
\end{itemize}
\end{lemma}

@@

\begin{lemma}
Posit $K=2$ arms and Bernoulli rewards.
\begin{itemize}
\item[(a)] Algorithm \PhasedExplorExploit is monotone.
\ascomment{Steven: for every \MRV or only in expectation?}
